---
title: "Week 8: Repeated Measures ANOVA"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
knitr::opts_chunk$set(echo = FALSE)
gradethis::gradethis_setup()
library(datarium)
library(tidyverse)
library(ez)
library(psyntur)
df <- selfesteem %>%
  gather(key = "time", value = "score", t1, t2, t3)
df <- df %>% 
  mutate_at(vars(id, time), as.factor)
df2 <- selfesteem2 %>%
  gather(key = "time", value = "score", t1, t2, t3)
df2 <- df2 %>% 
  mutate_at(vars(id, treatment, time), as.factor)
df2b <- df2 %>% 
mutate(df2, IV = interaction(time, treatment))
```

## Introduction

This is a self-assessment on repeated measures ANOVA, whether one-way or factorial ANOVA. Before you attempt this, you should first watch all the video lectures for Week 8, and also do the in-class workshop in Week 8.

This problem set will use the [selfeesteem](https://rdrr.io/cran/datarium/man/selfesteem.html) and the [selfeesteem2](https://rdrr.io/cran/datarium/man/selfesteem2.html) data sets, which we have used in the workshop.


## Explore data and test for assumptions

Let's start with one-way repeated measures ANOVA. Assume you were to test a null hypothesis of no difference between mean self-esteem scores, across three time points, as effect of a new type of diet.

Data (```df```) has already been loaded, in long format. Take a minute to explore it.

```{r explore, exercise=TRUE}
head(df)
```

Now test whether self-esteem scores within each time point are normally distributed, by using the Shapiro-Wilk test. Please run the ```shapiro_test``` function from the ```psyntur``` package on the ```df``` data frame. Packages are already loaded.

```{r norm, exercise=TRUE, exercise.eval=TRUE}

```

```{r norm-hint}
shapiro_test(score, by = time, data = df)
```

```{r norm-solution}
shapiro_test(score, by = time, data = df)
```
 
```{r norm-code-check}
grade_code()
```

```{r quiz0}
quiz(
  question("Based on the results from the Shapiro-Wilk test, which of the following is true?",
    answer("Data from t1 and t2 are normally distributed"),
    answer("Data from t2 and t3 are normally distributed"),
    answer("Data from t1 and t3 are normally distributed"),
    answer("Data from t1, t2 and t3 are normally distributed", correct = TRUE)
  )
)
```


## One-way repeated measures ANOVA

Now run a one-way ANOVA on the newly created ```df``` data and the ```ezANOVA``` function. Feel free to explore the data again, if you need. Packages are already loaded.

```{r one-way_anova, exercise=TRUE, exercise.eval=TRUE}

```

```{r one-way_anova-hint}
ezANOVA(data = df, dv = score, wid = id, within = time)
```

```{r one-way_anova-solution}
ezANOVA(data = df, dv = score, wid = id, within = time)
```
 
```{r one-way_anova-code-check}
grade_code()
```


### Quiz

Please now answer the following questions:

```{r quiz1}
quiz(
  question("Based on the results from the sphericity test, which of the following outputs would be best to interpret?",
    answer("Greenhouse-Geisser correction"),
    answer("Huynd-Feldt correction"),
    answer("Non-corrected solution", correct = TRUE)
  ),
  question("How would you interpret the main effect of diet on self-esteem?",
    answer("Non-significant"),
    answer("Significant at p < 0.001", correct = TRUE),
    answer("Significant at p < 0.01"),
    answer("Significant at p < 0.05")
  )
)
```


## Factorial repeated measures ANOVA

Let's now move on and run a factorial repeated measures ANOVA to test a null hypothesis of no difference between mean self-esteem scores, between two consecutive trials (diet, control) and three time points (t1, t2, t3).

Data was loaded in long format, take a minute to explore it. Packages are already loaded.

```{r explore2, exercise=TRUE}
head(df2)
```

Now plot self-esteem scores by ```treatment``` and ```time```, the data using the ```tukeyboxplot``` function from the ```psyntur``` package. Remember to use the ```df2``` data frame. 

```{r explore2x, exercise=TRUE, exercise.eval=TRUE}

```

```{r explore2x-hint}
tukeyboxplot(data = df2, y = score, x = time, by = treatment)
```

```{r explore2x-solution}
tukeyboxplot(data = df2, y = score, x = time, by = treatment)
```
 
```{r explore2x-code-check}
grade_code()
```

```{r quiz2}
quiz(
  question("Which of the following is correct?",
    answer("The lowest median self-esteem score was observed in the treatment group, t1"),
    answer("The lowest median self-esteem score was observed in the control group, t1"),
    answer("The lowest median self-esteem score was observed in the treatment group, t3"),
    answer("The lowest median self-esteem score was observed in the control group, t3", correct = TRUE)
  )
)
```


## Testing the factorial ANOVA model

Now let's run two-way ANOVA on the newly created ```df2``` data frame, using the ```ezANOVA``` function.

```{r two-way_anova, exercise=TRUE, exercise.eval=TRUE}

```

```{r two-way_anova-hint}
ezANOVA(data = df2, dv = score, wid = id, within = c(time, treatment))
```

```{r two-way_anova-solution}
ezANOVA(data = df2, dv = score, wid = id, within = c(time, treatment))
```
 
```{r two-way_anova-code-check}
grade_code()
```


### Quiz

Please now answer the following questions:

```{r quiz3}
quiz(
  question("Based on the results from the sphericity test, which of the following outputs would be most appropriate to use, to interpret the interaction?",
    answer("Greenhouse-Geisser correction"),
    answer("Huynd-Feldt correction"),
    answer("Non-corrected solution", correct = TRUE)
  ),
  question("How would you interpret the interaction of time and treatment on self-esteem?",
    answer("Non-significant"),
    answer("Significant at p < 0.001", correct = TRUE),
    answer("Significant at p < 0.01"),
    answer("Significant at p < 0.05")
  )
)
```

Now you can test for pairwise comparisons across the levels of the IVs Complete the code below by using the ```pairwise.t.test``` function on the ```df2b``` data frame. This is equivalent to the same data used to run the model, but with the addition of a new variable named ```IV```, combining all the levels of the independent variables. Please remember to correct for the family-wise error rate. 

```{r explore4, exercise=TRUE}
head(df2b)
```

```{r two-way_t_anova, exercise=TRUE, exercise.eval=TRUE}

```

```{r two-way_t_anova-hint}
pairwise.t.test(df2b$score, df2b$IV, paired = TRUE, p.adjust.method = "bonferroni")
```

```{r two-way_t_anova-solution}
pairwise.t.test(df2b$score, df2b$IV, paired = TRUE, p.adjust.method = "bonferroni")
```
 
```{r two-way_t_anova-code-check}
grade_code()
```
